Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-06-28 paper/paper.tex
Paper WIP

New introduction: Simpson and Goodhill presented a model in which the
chemoaffinity mechanism is non-biological. We offer a model in which a
competition mechanism counteracts the drive of axons along the ephrin
gradient. Will we need the axon-axon interaction mechanism based on
non-same-Eph repulsion?

Another approach: We present a model of gradient-following based on graded
interactions between the growth cones of retinal ganglion cell axons and the
tectal surface. We show that a combination of exponentially graded expression
in the retina and logarithmically graded expression on the tectum permits a
purely local model to reproduce the majority of reported phenomena associated
with retinotectal organisation.

@@

Our model consists of agents, representing axonal growth cones, originating
from a square retina and moving upon a square tectum. Each agent carries a set
of 4 receptors and 4 ligands which it expresses at differing levels depending
upon the soma's origin on the retina. The expression levels of
the receptors and ligands form gradients across the retinal surface. Receptors
are arranged in orthogonal pairs with $r_0$ being orthogonal to $r_1$ and
$r_2$, whose gradient is opposite to $r_0$ is orthogonal to $r_3$. Ligands are
also arranged in orthogonal pairs, with...
%
The tectum expresses the same sets of receptors and ligands, also in
orthogonal pairs of gradients, though not necessarily with the same expression
function.

>>>>

\textbf{New introduction:} Simpson and Goodhill presented a model in which the
chemoaffinity mechanism is non-biological. We offer a model in which a
competition mechanism counteracts the drive of axons along the ephrin
gradient. Will we need the axon-axon interaction mechanism based on
non-same-Eph repulsion?

\textbf{Another approach:} We present a model of gradient-following based on graded
interactions between the growth cones of retinal ganglion cell axons and the
tectal surface. We show that a combination of exponentially graded expression
in the retina and logarithmically graded expression on the tectum permits a
purely local model to reproduce the majority of reported phenomena associated
with retinotectal organisation.

@@

Our model consists of agents, representing axonal growth cones, originating
from a square retina and moving upon a square tectum. Each agent carries a set
of 4 receptors, $r$, and 4 ligands, $l$, which it expresses at differing
levels depending upon the soma's origin on the retina. The expression levels
of the receptors and ligands form gradients across the retinal
surface. Receptors are arranged in orthogonal pairs with the gradient of $r_0$
being orthogonal to that of $r_1$. $r_2$, whose gradient is opposite to $r_0$
is orthogonal to $r_3$. Ligands are also arranged in orthogonal pairs, with
$l_0$ opposing $r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$
orthogonal to $l_0$ and opposing $r_1$. While we do not explicitly name $r_0$,
$r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes
EphA, EphB and Ryk~\cite{schmitt_wntryk_2006} receptors and that $l$ includes
the ephrin-A, ephrin-B and Wnt3~\cite{schmitt_wntryk_2006} ligands.
%
The tectum expresses the same sets of receptors and ligands, also in
orthogonal pairs of gradients, though not necessarily with the same expression
function. Figure~\ref{f:1} shows colour maps of receptor and ligand
expression, in this case experssion maps are linear.
